\documentclass[10pt]{article}
%imports extra math symbols
\usepackage{amssymb}
%used for doublespace text
\usepackage{setspace}
% uncomment the next line to use graphics
%\usepackage{graphicx}
%shortnames for common commands
\newcommand{\BD}{\begin{doublespace}}
\newcommand{\ED}{\end{doublespace}}
\newcommand{\BE}{\begin{enumerate}}
\newcommand{\EE}{\end{enumerate}}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\BT}{\begin{tabular}}
\newcommand{\ET}{\end{tabular}}
\newcommand{\BV}{\begin{verbatim}}
\newcommand{\EV}{\end{verbatim}}
\newcommand{\BM}{\begin{displaymath}}
\newcommand{\EM}{\end{displaymath}}
\newcommand{\BC}{\begin{center}}
\newcommand{\EC}{\end{center}}
\newcommand{\BFR}{\begin{flushright}}
\newcommand{\EFR}{\end{flushright}}
\newcommand{\BFL}{\begin{flushleft}}
\newcommand{\EFL}{\end{flushleft}}
%Stretches out the page
\setlength{\oddsidemargin}{-0.50in}
\setlength{\evensidemargin}{-0.50in}
\setlength{\textwidth}{7.0in}
\setlength{\topmargin}{0.0in}
\setlength{\textheight}{9.0in}
\setlength{\parindent}{0cm}
\setlength{\parskip}{0ex}
\pagestyle{empty}
%Sets title page information
\author{Brandon Chisham}
\title{Queries of Interest}
\date{\today}
\begin{document}
%Generates the title page
\maketitle
%uncomment the next line to generate a table of contents
%\tableofcontents

\section{Bibe03 Paper}
 \BE
   \item  Given a set \textit{S} of taxa, find a minimum spanning clade
   \item  Given three sets of taxa $S_1$, $S_2$, and $S_3$  and based
             on all the phylogenies in the DB find the relationship between the
             specified taxa.
   \item  Given a set of taxa $S$ return all the phylogenies that contain $S$.
   \item  Given a phylogenetic method $M$ find all phylogenies that were
             reconstructed using $M$.
   \item: Given an integer $n$, representing a number of taxa find all phylogenies
             on a set of taxa $S$ such that $\vert S \vert \approx F(n).$
   \item: Retrieve phylogenies based on date or author.
   \item  Given some characteristic $l$ find phylogenies with that characteristic.
             Characteristics include shortest edge, longest edge, tree diameter, and tree width.
   \item  Given a datatype return all trees built using a dataset of that type.
   \item  Given a phylogeny $T$ and a measure $M$ and a quality $c$ find all phylogenies that are
             $c$ units from $T$ using $M$.
   \item  Return all phylogenies created using some specified model of evolution.
   \item  Given some measures return stats about the database with regard to the specified measures. 
 \EE

\section{VoCamp}
 \BE
  \item Linked Data. At the meeting a variety of the use cases involved linking data from separate sources 
        especially relating organisms with some character state to a particular place. Or given a particular
        place return the organisms found there, along with the states of some characters of interest.
  \item Given a set of taxa find the MRCA of them.
 \EE

\section{VoCamp Followup Conference Call}
 \BE
  \item Given a set of organisms defining a clade return all subtrees containing the organisms in the set.
 \EE
\end{document}

